A Formal Model of Checked C

### 2023-04-08 sandp23/background.tex
Finishing up background

\section{Background and Motivation}
This section presents the necessary background and motivation for our work.
\aravind{I am working on this.}
%\section{Overview and Transcendence}\label{sec:overview}
%This section describes \checkedc{} and new features \systemname{} provides.

@@

\subsection{Program Partioning}
\label{subsec:background:programpart}
Program partitioning~\cite{rul2009towards} is a well-known technique for dividing a program into multiple isolated parts or partitions.
There has been considerable work~\cite{tan2017principles, brumley2004privtrans, bittau2008wedge, lind2017glamdring, liu2017ptrsplit} in the area.
Most of these techniques are~\emph{data-centric}~\cite{lind2017glamdring, liu2017ptrsplit}, wherein program data drives the partitioning.
Specifically, given sensitive data in a program, all functions accessing sensitive data will be considered part of the trusted partition, and all other functions will be in the untrusted partition.
For instance, in the following code~\code{key} is marked as sensitive and functions that access it,~\ie{}~\code{initkey} and~\code{encrypt} are considered to be part of trusted parition and the other function,~\ie{}~\code{main} will be in the untrusted partition.
\begin{minted}[xleftmargin=25pt, mathescape, linenos, escapeinside=||, fontsize=\footnotesize{}]{c}
char __attribute__((annotate("sensitive"))) *key;

@@

As shown by the recent study from Microsoft, 70\% of the vulnerabilities still occur due to spatial safety violations.
Several works show that certain functions are more prone to have a certain class of vulnerabilities than others.
Moreover, certain vulnerabilities may be impossible to occur in a given function. For instance, a function that cannot invoke (directly or transitively) any heap operations (\eg{} malloc/free) cannot contain temporal violations,~\eg use-after-free.
Instead of complete isolation, we need~\emph{\acf{SeFI}} from a program partition.
Formally, we say that a partition is $\phi$-\ac{SeFI} if and only if any violation of the property $\phi$ is contained with the partition and does not affect other partition(s).
For instance,~\emph{spatial safety-\ac{SeFI}} from a partition provides isolation against spatial safety violations in the partition.

@@

Although the blameless proof exists~\cite{ruef2019achieving, li22checkedc} for~\checkedc, it does not state that spatial safety violations cannot happen in checked regions but rather states that checked regions~\emph{cannot be blamed for any spatial safety violations}.
Consider the following example:
\begin{minted}[xleftmargin=25pt, mathescape, linenos, escapeinside=||, fontsize=\footnotesize{}]{c}
// c region code
int func(array_ptr<char> p : count(5)) {
|\textcolor{red}{\faChainBroken}|..p[4]..
}
// u region code
...
str = "he";
...
func(|\textcolor{red}{\faBug}|assume_bounds_cast<char>(str, 5)); 
\end{minted}
Here, the checked function~\inlinecode{func} expects a pointer to a buffer of five elements, but the unchecked region code invokes the function (Line 9) with a buffer of 2 elements.
This results in a spatial safety violation (\textcolor{red}{\faChainBroken}) in the checked function, but of course, the blame or the root cause is in the unchecked function (\textcolor{red}{\faBug}).
Furthermore, since checked and unchecked functions execute in the same address space, spatial memory corruptions (\eg buffer overflow) in unchecked functions (or regions) can take down the complete program despite
having checked regions.
We need~\emph{an isolation mechanism to ensure that code executed as part of unchecked functions does not violate the safety guarantees in checked functions}.

>>>>

\section{Background and Motivation}
This section presents the necessary background and motivation for our work.
%\section{Overview and Transcendence}\label{sec:overview}
%This section describes \checkedc{} and new features \systemname{} provides.

@@

\subsection{Program Partioning}
\label{subsec:background:programpart}
Program partitioning~\cite{rul2009towards} is a well-known technique for dividing a program into multiple isolated parts or partitions.
There has been considerable work~\cite{tan2017principles, brumley2004privtrans, bittau2008wedge, lind2017glamdring, liu2017ptrsplit} in the area.
Most of these techniques are~\emph{data-centric}~\cite{lind2017glamdring, liu2017ptrsplit}, wherein program data drives the partitioning.
Specifically, given sensitive data in a program, all functions accessing sensitive data will be considered part of the trusted partition, and all other functions will be in the untrusted partition.
For instance, in the following code~\code{key} is marked as sensitive and functions that access it,~\ie{}~\code{initkey} and~\code{encrypt} are considered to be part of trusted parition and the other function,~\ie{}~\code{main} will be in the untrusted partition.
\begin{minted}[xleftmargin=25pt, mathescape, linenos, escapeinside=||, fontsize=\scriptsize{}]{c}
char __attribute__((annotate("sensitive"))) *key;

@@

As shown by the recent study from Microsoft~\cite{microsoftmemsafe}, 70\% of the vulnerabilities still occur due to spatial safety violations.
Several works~\cite{meng2021bran, du2019leopard, ding2022velvet, situ2021vulnerable} show that certain functions are more prone to have a certain class of vulnerabilities than others.
Moreover, certain vulnerabilities may be impossible to occur in a given function. For instance, a function that cannot invoke (directly or transitively) any heap operations (\eg{} malloc/free) cannot contain temporal violations,~\eg use-after-free.
Instead of complete isolation, we need~\emph{\acf{SeFI}} from a program partition.
Formally, we say that a partition is $\phi$-\ac{SeFI} if and only if any violation of the property $\phi$ is contained with the partition and does not affect other partition(s).
For instance,~\emph{spatial safety-\ac{SeFI}} from a partition provides isolation against spatial safety violations in the partition.

@@

Although the blameless proof exists~\cite{ruef2019achieving, li22checkedc} for~\checkedc, it does not state that spatial safety violations cannot happen in checked regions but rather states that checked regions~\emph{cannot be blamed for any spatial safety violations}.
Consider the following example:
\begin{minted}[xleftmargin=25pt, mathescape, linenos, escapeinside=||, fontsize=\scriptsize{}]{c}
// code with checked c annotations
int func(array_ptr<char> p : count(5)) {
|\textcolor{red}{\faChainBroken}|..p[4]..
}
// unchecked (wild) code
...
str = "he";
...
func(|\textcolor{red}{\faBug}|assume_bounds_cast<char>(str, 5)); 
\end{minted}
Here, the checked function~\inlinecode{func} expects a pointer to a buffer of five elements, but the unchecked region code invokes the function (Line 9) with a buffer of 2 elements.
This results in a spatial safety violation (\textcolor{red}{\faChainBroken}) in the checked function, but of course, the blame or the root cause is in the unchecked function (\textcolor{red}{\faBug}).
Furthermore, since checked and unchecked functions execute in the same address space, spatial memory corruptions (\eg buffer overflow) in unchecked functions (or regions) can take down the complete program despite
having checked regions.
We need~\emph{an isolation mechanism to ensure that code executed as part of unchecked functions does not violate the safety guarantees in checked functions}.
